---
title: "Sequoia"
author: "Sarah P. Flanagan"
date: "May 7, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="C:/Users/sflan/Documents/GitHub/SCA/results/")
```
```{r}
library(sequoia)
```

## Setup

First I need to get the genomic data in the correct format. 

And create ped and map files so that I can easily prune the loci

```{r}
gt<-read.delim("relatedness/coancestry_gty.txt",sep='\t',header=FALSE)
ped<-data.frame(FamID=as.character(LifeHistData$ID),IndID=as.character(LifeHistData$ID),PatID=0,
                MatID=0,Sex=LifeHistData$Sex,Phenotype=-9,
                gt[,2:ncol(gt)],stringsAsFactors = FALSE)
ped[grep("OF",ped[,1]),1]<-gsub("OF_(\\d+)(-\\d+)?","\\1",ped[grep("OF",ped[,1]),1])
ped[grep("PR",ped[,1]),1]<-gsub("PR_(\\d+)","\\1",ped[grep("PR",ped[,1]),1])
ped[ped[,1]=="086-23",1]<-"08623"
ped[ped[,1]=="086R",1]<-"086"
ped[,3]<-ped[,1]
write.table(ped,"sequoia/GenoM.ped",sep='\t',row.names=FALSE,col.names=FALSE,quote=FALSE)
#create map file
vcf<-read.table("relatedness/drad_miss3.vcf",header=TRUE)
map<-data.frame(vcf$CHROM,vcf$ID,NA,vcf$POS)
write.table(map,"sequoia/GenoM.map",sep='\t',row.names=FALSE,col.names=FALSE,quote=FALSE)
```

`plink --file C:\Users\sflan\Documents\GitHub\SCA\results\sequoia\GenoM --maf 0.4 --indep 50 5 2 --allow-extra-chr --make-founders --recodeA --out C:\Users\sflan\Documents\GitHub\SCA\results\sequoia\PrunedGenoM` for raw input or
`plink --file C:\Users\sflan\Documents\GitHub\SCA\results\sequoia\GenoM --maf 0.4 --indep 50 5 2 --allow-extra-chr --make-founders --recode --out C:\Users\sflan\Documents\GitHub\SCA\results\sequoia\PrunedGenoM` for ped input

This resulted in 2312 loci

```{r}

coan2sequoia<-function(gt.name,sep='\t'){
  gt<-read.delim(gt.name,sep=sep,header=FALSE)
  GenoM<-as.matrix(do.call(cbind,lapply(seq(2,ncol(gt),2),function(coln,gt){
    dat<-gt[,c(coln,coln+1)]
    ref<-names(which.max(table(apply(dat,2,factor))))
    out<-apply(dat,1,function(ind,ref){
      if(ind[1]==ind[2]){
        if(ind[2]==ref){ out<-2 }
        if(ind[2]==0){ out<--9 }
        if(ind[2]!=ref & ind[2] != 0 ){ out<-0 }
      }else{
        out<-1
      }
      return(out)
    },ref)
  },gt)))
  rownames(GenoM)<-gt[,1]
  return(GenoM)
}
GenoM<-coan2sequoia("relatedness/coancestry_gty.txt")
summary(SnpStats(GenoM))

ped2sequoia<-function(gt.name,sep='\t'){
  gt<-read.delim(gt.name,sep=sep,header=FALSE)
  GenoM<-as.matrix(do.call(cbind,lapply(seq(7,ncol(gt),2),function(coln,gt){
    dat<-gt[,c(coln,coln+1)]
    ref<-names(which.max(table(apply(dat,2,factor))))
    out<-apply(dat,1,function(ind,ref){
      if(ind[1]==ind[2]){
        if(ind[2]==ref){ out<-2 }
        if(ind[2]==0){ out<--9 }
        if(ind[2]!=ref & ind[2] != 0 ){ out<-0 }
      }else{
        out<-1
      }
      return(out)
    },ref)
  },gt)))
  rownames(GenoM)<-gt[,1]
  return(GenoM)
}
GenoM_ped<-ped2sequoia("sequoia/PrunedGenoM.ped",sep=' ')
summary(SnpStats(GenoM_ped))

GenoM_ped<-GenoConvert(InFile = "sequoia/PrunedGenoM.raw")
```


Then I need to generate the life history data

```{r}
sex<-gsub(".*(\\w{2})_.*","\\1",rownames(GenoM_ped))
sex[sex=="FE"]<-as.numeric(1)
sex[sex%in%c("NP","PR")]<-as.numeric(2)
sex[sex=="OF"]<-NA
BY<-sex
BY[BY %in% c("1","2")]<-1
BY[is.na(BY)]<-2
LifeHistData<-data.frame(ID=rownames(GenoM_ped),Sex=as.numeric(sex),BY=as.numeric(BY))
```



## Run Sequoia

Now I'm ready to run sequoia. `MaxSibIter` performs sibship clustering and is time consuming - wise to run with `MaxSibIter=0` for only parentage assignment. It's unclear how to set `Tassign` or `Tfilter`. 

```{r}
seq<-sequoia(GenoM,LifeHistData,MaxSibIter=0,MaxSibshipSize = 200)
seq_ped<-sequoia(GenoM_ped,LifeHistData,MaxSibIter=0,MaxSibshipSize = 200,Tassign = 0,MaxMismatch = 150)
```

For some reason these fail to assign parentage.